Mixing Time Estimation in Reversible Markov Chains from a Single Sample Path

### 2015-11-03 arxiv-v2/lowerbound.tex
typo
.

  In order to guarantee $|\hatgap - \gap| < 1/8 < |1 -
  (1/2+\bar\pi)|/2$, it must be possible to distinguish the two Markov
  chains.
  Assume that the initial state distribution has mass at least $1/2$
  on state $1$.
  (If this is not the case, we swap the roles of states $1$ and $2$ in
  the constructions above.)
  With probability at least half, the initial state is $1$; and both
  chains have the same transition probabilities from state $1$.
  The chains are indistinguishable less the sample path eventually
  reaches state $2$.
  But with probability at least $3/4$, a sample path of length $n <
  1/(4\bar\pi)$ starting from state $1$ always remains in the same
  state (this follows from properties of the geometric distribution
  and the assumption $\bar\pi < 1/4$).
\end{proof}

@@

We claim in Section~\ref{sec:intro} that a sample path of length
$\Omega\parens{ (d \log d)/\gap + 1/\pimin}$ is required to guarantee
constant multiplicative accuracy in estimating $\gap$.
This follows by combining Theorems~\ref{thm:lb-pimin}
and~\ref{thm:lb-gap} in a standard, straightforward way.
Specifically, if the length of the sample path $n$ is smaller than
$(n_1 + n_2) / 2$---where $n_1$ is the lower bound from
Theorem~\ref{thm:lb-pimin}, and $n_2$ is the lower bound from
Theorem~\ref{thm:lb-gap}---then $n$ is smaller than
$\max\braces{n_1,n_2}$.
So at least one of Theorem~\ref{thm:lb-pimin} and
Theorem~\ref{thm:lb-gap} implies the existence of an ergodic and
reversible Markov chain distribution with (say) $1/4 \leq \gap \leq
1/2$ and $\pimin \geq 1/(2d)$ such that
\[
  \Pr\Brackets{ |\hatgap - \gap| \geq \gap/4 } \geq 1/4
  .
\]

>>>>

  In order to guarantee $|\hatgap - \gap| < 1/8 < |1 -
  (1/2+\bar\pi)|/2$, it must be possible to distinguish the two Markov
  chains.
  Assume that the initial state distribution has mass at least $1/2$
  on state $1$.
  (If this is not the case, we swap the roles of states $1$ and $2$ in
  the constructions above.)
  With probability at least half, the initial state is $1$; and both
  chains have the same transition probabilities from state $1$.
  The chains are indistinguishable unless the sample path eventually
  reaches state $2$.
  But with probability at least $3/4$, a sample path of length $n <
  1/(4\bar\pi)$ starting from state $1$ always remains in the same
  state (this follows from properties of the geometric distribution
  and the assumption $\bar\pi < 1/4$).
\end{proof}

@@

We claim in Section~\ref{sec:intro} that a sample path of length
$\Omega\parens{ (d \log d)/\gap + 1/\pimin}$ is required to guarantee
constant multiplicative accuracy in estimating $\gap$.
This follows by combining Theorems~\ref{thm:lb-pimin}
and~\ref{thm:lb-gap} in a standard, straightforward way.
Specifically, if the length of the sample path $n$ is smaller than
$(n_1 + n_2) / 2$---where $n_1$ is the lower bound from
Theorem~\ref{thm:lb-pimin}, and $n_2$ is the lower bound from
Theorem~\ref{thm:lb-gap}---then $n$ is smaller than
$\max\braces{n_1,n_2}$.
So at least one of Theorem~\ref{thm:lb-pimin} and
Theorem~\ref{thm:lb-gap} implies the existence of an ergodic and
reversible Markov chain distribution (with $\pimin \geq 1/(2d)$) such that
\[
  \Pr\Brackets{ |\hatgap - \gap| \geq \gap/4 } \geq 1/4
  .
\]
